An Efficient Calculus for Generalised Parikh Images of Regular Languages

### 2022-07-06 main.tex
started rewriting Ostrich+ section

\begin{figure}[t]
  \begin{minipage}[b]{0.50\linewidth}
  \centering 
    \includegraphics[width=\textwidth]{aca_or_bc}
    \caption{An automaton recognising the regular expression
    $\mathtt{ac^*a|b(bb)^*c}$.}\label{fig:aca-or-bc}

    \includegraphics[width=\textwidth]{something_c_something}
    \caption{An automaton recognising the regular expression
    $\mathtt{.^*c.^*}$.}\label{fig:something-c-something}
  \end{minipage}
  \Description[Two deterministic finite automata, one above the other]{NFAs 
    recognising the languages ac^*a|b(bb)^*c and ac^*a|b(bb)^*c respectively.
    The first one has 4 states and the second one 2.}
  \end{figure}

@@

\subsection{Example~2: Generalised String Constraints with Integer
  Datatype}\label{sec:parikh-automata}
\contents{\item Ostrich+!}

Another application for this generalisation is to express Parikh automata, and
this is indeed what the implementation described in
Section~\ref{sec:implementation} does. Following~\cite{parikh-automata} as
filtered through~\cite{expressiveness} (which offers a better formalisation for
our use), we define a Parikh automaton as an automaton with integer counters
that are incremented at the transitions, and where we have linear
constraints on the final values of the counters after running the automaton, or
formally as in \cref{def:parikh-automata,def:projections}.

@@

\begin{figure}[t]
  \centering
      \includegraphics[width=0.5\linewidth]{parikh_automaton}
      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
      formula containing the constraints on the final register values cannot be
      visualised.}\label{fig:parikh-automaton}
    \end{figure}

>>>>

\begin{figure}[t]
  \begin{minipage}[b]{0.70\linewidth}
  \centering 
  \includegraphics[scale=0.9]{aca_or_bc}
  \caption{An automaton recognising the regular expression
    $\mathtt{ac^*a|b(bb)^*c}$. \Fudge{this is actually not the right
      regular expression. the automaton also accepts words in which a
      b is followed by a.}}\label{fig:aca-or-bc}

    \includegraphics[scale=0.9]{something_c_something}
    \caption{An automaton recognising the regular expression
    $\mathtt{.^*c.^*}$.}\label{fig:something-c-something}
  \end{minipage}
  \Description[Two deterministic finite automata, one above the other]{NFAs 
    recognising the languages ac^*a|b(bb)^*c and ac^*a|b(bb)^*c respectively.
    The first one has 4 states and the second one 2.}
  \end{figure}

@@

\subsection{Example~2: Generalised String Constraints with Integer
  Datatype}\label{sec:parikh-automata}

Parikh images are also applicable to decide more general classes of
string constraints~\cite{ostrich-plus}. Consider, for instance, a
constraint involving the substring operation:
\begin{equation}
  \label{eq:substring}
  x \in \Language_1 \wedge \; 0 \leq n \leq m < |x| \wedge x[n:m] \in \Language_2
\end{equation}
in which $x$ is a string variable, $n, m$ are integer variables, and
$x[n:m]$ denotes the substring of $x$ starting at position~$n$ running
until position~$m$. To illustrate the decision procedure in
\cite{ostrich-plus} for string constraints of this kind, suppose that
$\Language_2$ is defined by the regular
expression~$\mathtt{(ab|c)*}$. An automaton recognising this language
is shown in Fig.~\ref{fig:parikh-automata} (left).

We can model \eqref{eq:substring} using the notion of a \emph{Parikh
  automaton}~\cite{parikh-automata,expressiveness}. A Parikh automaton
is an automaton in which transitions are labelled both with characters
from the alphabet~$\Alphabet$, and with offset vectors defining the
increments of a finite number of counters.

@@

\iffalse
\begin{figure}[t]
  \centering
      \includegraphics[width=0.5\linewidth]{parikh_automaton}
      \caption{The automaton part of a Parikh automaton for $\AcaOrBc{}$ with
      $\Alphabet = \Set{\text{a, b, c}}, d = 3$. The semilinear set/Presburger
      formula containing the constraints on the final register values cannot be
      visualised.}\label{fig:parikh-automaton}
    \end{figure}
\fi

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.9]{counter_automaton_pre}
  \hfill
  \includegraphics[scale=0.9]{counter_automaton}
  \caption{An automaton recognising the language $\mathtt{(ab|c)*}$
    (left), and a Parikh automaton describing the pre-image of the
    substring operation (right).}\label{fig:parikh-automata}
    \end{figure}
